---
title: "SSC 442 Lab 01"
author: "Team 7"
date: "1/16/2020"
output: pdf_document
---

```{r include=FALSE}
library(tidyverse)
library(knitr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
mpgData <- ggplot2::mpg
bankData <- read.csv("bank.csv")
```
# **Exercise 1:**

```{r dispVHwy}
DispVHwy <- ggplot(
  data=mpgData,
  mapping=aes(
    x=displ,
    y=hwy
  )
) + 
  geom_point() +
  xlab("Engine Size (Liters)") + 
  ylab("Highway Fuel Efficiency (mpg)") + 
  ggtitle("Engine Size Vs. Highway Fuel Efficiency") +
  theme(plot.title=element_text(hjust=0.5))

DispVHwy
```

This scatter plot of Engine Size vs. Highway Fuel Efficiency shows a negative correlation between engine size and highway fuel efficiency. That is, the scatter plot demonstrates a general trend that as engine size (disp) increases, highway fuel efficiency will decrease. This relationship is, however, intuitive. One would expect that as the size of the engine increases, the engine will use more gasoline, and thus will have a lower highway fuel efficiency.

```{r classVDrv}
ClassVDrv <- ggplot(
  data=mpgData,
  mapping=aes(
    x=class,
    y=drv
  )
) + 
  geom_point() + 
  xlab("Vehicle Class") + 
  ylab("Vehicle Drive Type") +
  ggtitle("Vehicle Class vs. Drive Type") +
  theme(plot.title=element_text(hjust=0.5))

ClassVDrv
```

This scatter plot of Vehicle Class Vs. Drive Type is not a useful plot because there is no related relationship between
the two variables. The vehicle class generally dictates what drive type the vehicle may have, and there are far too few data points to make any reasonable conclusions

# **Exercise 1b:**

```{r classVDrvColor}
DisplVHwyColor <- ggplot(
  data=mpgData,
  mapping=aes(
    x=displ,
    y=hwy,
    color=class
  )
) + 
  geom_point() +
  xlab("Engine Size (Liters)") + 
  ylab("Highway Fuel Efficiency (mpg)") + 
  ggtitle("Engine Size Vs. Highway Fuel Efficiency, Colored by Vehicle Class")

DisplVHwyColor
```

Given the scatter plot Engine Size Vs. Highway Fuel Efficiency generated with color coded dots according to the type
of vehicle, it can be noted that 2-seat vehicles are outlyers in the sense that they achieve better highway fuel efficiency than other types of vehicles with the same size engine. This is likely explained by the fact that 2-seat vehicles are smaller and lighter than the other vehicles which have similar sized engines.

# **Exercise 2:**

```{r, results='asis'}
MaritalDefaultPlot <- ggplot(
  data=bankData,
  mapping=aes(
    x=marital,
    fill=default,
  )
) +
  geom_bar(position="dodge") +
  xlab("Marital Status") + 
  ylab("Number of Customers") + 
  scale_fill_manual(
    values=c("black", "red")
  ) + 
  ggtitle("Marital Status and Default Rates with Education Levels") +
  theme(plot.title=element_text(hjust=0.5))

MaritalDefaultPlot

numDivorced <- length(which(bankData$marital == "divorced"))
numMarried <- length(which(bankData$marital == "married"))
numSingle <- length(which(bankData$marital == "single"))
numTotal <- numDivorced + numMarried + numSingle

numDivorcedDefault <- length(bankData[bankData$marital=="divorced" & bankData$default=="yes", ]$default)
numMarriedDefault <- length(bankData[bankData$marital=="married" & bankData$default=="yes", ]$default)
numSingleDefault <- length(bankData[bankData$marital=="single" & bankData$default=="yes", ]$default)
numTotalDefault <- numDivorcedDefault + numMarriedDefault + numSingleDefault

totalNumCustomers <- c(numDivorced, numMarried, numSingle, numTotal)
totalNumCustomersDefaulted <- c(numDivorcedDefault, numMarriedDefault, numSingleDefault, numTotalDefault)
percentageDefaulted <- c(
  ((numDivorcedDefault  /numDivorced) * 100), 
  ((numMarriedDefault / numMarried) * 100), 
  ((numSingleDefault / numSingle) * 100),
  ((numTotalDefault / numTotal) * 100)
)

MaritalStatusDefaultAnalysis <- as.data.frame(totalNumCustomers, c("Divorced", "Married", "Single", "Total"))
MaritalStatusDefaultAnalysis$totalNumCustomersDefaulted <- totalNumCustomersDefaulted
MaritalStatusDefaultAnalysis$percentageDefaulted <- percentageDefaulted

kable(MaritalStatusDefaultAnalysis[,])
```

The above graph and table dmeonstrate how customer marital status is related to the rate at which they default on the loan or mortgage the customer holds with the bank. From the graph and table, we note that in general, a very small proportion (approximatley 1.7%) of all customers, across all marital stauses, default on the loans/mortgages they hold. However, the table demonstrates that customers who have been divorced are the most at-risk population for defaulting on their loan/mortgage, and married individuals are the population with the lowest risk of defaulting.

```{r, results='asis'}
MortgagesAndLoans <- ggplot(
  data=bankData,
  mapping=aes(
    x=housing,
    fill=loan
  )
) + 
  geom_bar(
    position=position_dodge()
  ) + 
  scale_fill_manual(
    values=c("red", "blue")
  ) + 
  xlab("Has a Mortgage") +
  ylab("Number") + 
  ggtitle("Mortgages and Loans") +
  theme(plot.title=element_text(hjust=0.5))

MortgagesAndLoans

numNeither <- length(bankData[bankData$housing!="yes" & bankData$loan!="yes", ]$housing)
numJustMortgage <- length(bankData[bankData$housing=="yes" & bankData$loan!="yes", ]$housing)
numJustLoan <- length(bankData[bankData$housing!="yes" & bankData$loan=="yes", ]$housing)
numBoth <- length(bankData[bankData$housing=="yes" & bankData$loan=="yes", ]$housing)
totalWithMortgageOrLoan <- numJustMortgage + numJustLoan + numBoth
total <- totalWithMortgageOrLoan + numNeither

MortgageLoanAnalysis <- as.data.frame(numNeither, "Number")
MortgageLoanAnalysis$a <- numJustMortgage
MortgageLoanAnalysis$b <- numJustLoan
MortgageLoanAnalysis$c <- numBoth

names(MortgageLoanAnalysis) <- c("Neither a Mortgage nor a Loan", "Only a Mortgage", "Only a Loan", "Both a Mortgage and a Loan")
MortgageLoanAnalysis <- rbind(
  MortgageLoanAnalysis,
  c(
    ((numNeither / total) * 100),
    ((numJustMortgage / total) * 100), 
    ((numJustLoan / total) * 100),
    ((numBoth / total) * 100)
  )
)
row.names(MortgageLoanAnalysis) <- c("Number of Customers", "Percentage")
MortgageLoanAnalysis <- t(MortgageLoanAnalysis)

kable(MortgageLoanAnalysis[,])
```

The above graph and table display data on customers who have neither a loan nor a loan, only a mortgage, only a loan, or both. From the graph, it can be seen that the largest population has only a mortgage, but the smallest population has both a loan and a mortgage. The table breaks these populations down into percentages, and it can be seen that roughly half of all customers hold mortgages, but the next largest population is customers who have neither a mortgage nor a loan. Thus, from a business perspective, one of the best approaches to increasing revenues for the bank would be to entice the large population of customers (approximately 37%) who do not have a mortgage or a loan to take out a mortgage or loan from the bank. Another tactic which could possibly increase business revenues would be to target the population of customers who have exclusively a loan or a mortgage, and attempt to entice them to take out another or one of the other types.